\begin{thebibliography}{10}

\bibitem{asu} Alfred V. Aho, Ravi Sethi and Jeffrey D. Ullman.
``Compilers - Principles, Techniques and Tools.'' Addison-Wesley, 1986.

\bibitem{beeskar} Kenneth R. Beesley and Lauri Karttunen. ``Finite-State 
Morphology: Xerox Tools and Techniques.'' Private communication, April 2001.

\bibitem{bentley} Jon L. Bentley and Robert Sedgewick.
``Fast Algorithms for Sorting and Searching Strings.'' 
Proceedings, 8th Annual ACM-SIAM Symposium on Discrete Algorithms, Jan. 1997.

\bibitem{berrysethi}
G\'erard Berry and Ravi Sethi.
From regular expressions to deterministic automata.
Theoretical Computer Science 48 (1986), pp. 117--126.

\bibitem{berstelpin}
Jean Berstel and Jean-Eric Pin. Local languages and the {Berry}-{Sethi} algorithm.
Theoretical Computer Science 155 (1996), pp. 439--446.

\bibitem{brill} Eric Brill. ``A simple rule-based part of speech tagger.''
In Proceedings, Third Conference on Applied Natural Language Processing, 1992.
Trento, Italy, 152--155. 

\bibitem{burge} W. H. Burge. ``Recursive Programming Techniques.''
Addison-Wesley, 1975. 

\bibitem{fp} Guy Cousineau and Michel Mauny. ``The Functional Approach to
Programming.'' Cambridge University Press, 1998.

\bibitem{daciuk} Jan Daciuk, Stoyan Mihov, Bruce W. Watson and Richard E. 
Watson. ``Incremental Construction of Minimal Acyclic Finite-State Automata.''
Computational Linguistics 26,1 (2000). 

\bibitem{eilenberg}
Samuel Eilenberg. Automata, Languages, and Machines, volume A.
Academic Press, 1974.

\bibitem{MLer} Matthias Felleisen and Daniel P. Friedman. ``The Little MLer''. 
MIT Press, 1998.

\bibitem{flajsipste} Philippe Flajolet, Paola Sipala and Jean-Marc Steyaert.
``Analytic Variations on the Common Subexpresssion Problem.'' Proceedings of
17th ICALP  Colloquium, Warwick (1990), LNCS 443, Springer-Verlag,
pp. 220--234.

\bibitem{ML-LCF}
M.  Gordon, R.  Milner, C.  Wadsworth.
``A Metalanguage for Interactive Proof in LCF.''
Internal Report CSR-16-77, Department of Computer Science,
University of Edinburgh (Sept. 1977).

\bibitem{zipper} G\'erard Huet. ``The Zipper''. J. Functional Programming 7,5 
(Sept. 1997), pp. 549--554.

\bibitem{dico-report} 
G\'erard Huet.
``Structure of a Sanskrit dictionary.''
INRIA Technical Report, Sept. 2000.
Available as: \raggedright
\verb!http://pauillac.inria.fr/~huet/PUBLIC/Dicostruct.ps!.

\bibitem{wcre}
G\'erard Huet.
``From an informal textual lexicon to a well-structured lexical database:
An experiment in data reverse engineering.''
IEEE Working Conference on Reverse Engineering (WCRE'2001), 
Stuttgart, Oct. 2001.

\bibitem{2003-Huet-3} 
G\'erard Huet. Automata Mista. In ``Verification: Theory and Practice: Essays Dedicated
to {Zohar} {Manna} on the Occasion of His 64th Birthday". Ed. Nachum Dershowitz,
Springer-Verlag LNCS vol. 2772 (2004), pp. 359--372.

\bibitem{2004-Huet-1}
G\'erard Huet. A Functional Toolkit for Morphological
and Phonological Processing, Application to a {Sanskrit} Tagger.
J. Functional Programming, 15,4 (2005), pp. 573--614. 

\bibitem{2006-Huet-Razet}
G\'erard Huet and Beno{\^\i}t Razet. The Reactive Engine for Modular Transducers.
In ``Algebra, Meaning and Computation, Essays Dedicated to 
Joseph A. Goguen on the Occasion of His 65th Birthday'', Eds.
Kokichi Futatsugi, Jean-Pierre Jouannaud and Jos\'e Meseguer.
Springer-Verlag LNCS vol. 4060 (2006), pp. 355--374

\bibitem{2008-Huet-Razet}
G\'erard Huet and Beno{\^\i}t Razet. Computing with Relational Machines.
ICON'2008 tutorial. Preliminary version available at URL
\url{http://yquem.inria.fr/~huet/PUBLIC/Pune_tutorial.pdf}.

\bibitem{kk} Ronald M. Kaplan and Martin Kay. ``Regular Models of
Phonological Rule Systems.'' Computational Linguistics (20,3), 1994,
pp. 331--378. 

\bibitem{karttunen1} Lauri Karttunen. ``Applications of Finite-State 
Transducers in Natural Language Processing.'' 
In Proceedings of CIAA-2000.

\bibitem{karttunen2} Lauri Karttunen. ``The Replace Operator.'' 
In Proceedings of ACL'95, Cambridge, MA, 1995. Extended version
in \cite{rs2}.

\bibitem{kosk} K. Koskenniemi. ``A general computational model for word-form
recognition and production.'' In Proceedings, 10th International Conference 
on Computational Linguistics, Stanford (1984). 

\bibitem{laporte} Eric Laporte. 
``Rational Transductions for Phonetic Conversion and Phonology.'' 
Report IGM 96-14, Institut Gaspard Monge, 
Universit\'e de Marne-la-Vall\'ee, Aug. 1995. Also in \cite{rs2}.

\bibitem{ocaml} Xavier Leroy et al. ``Objective Caml.'' See: \raggedright
\verb!http://caml.inria.fr/ocaml/index.html!.

\bibitem{mohri} Mehryar Mohri. ``Finite-State Transducers in Language and 
Speech Processing.'' Computational Linguistics 23,2 (1997), pp. 269--311.

\bibitem{paulson} Larry C. Paulson. ``ML for the Working Programmer.''
Cambridge University Press, 1991.

\bibitem{ranta} Aarne Ranta. ``The GF Language: Syntax and Type System.''
See: \raggedright \verb!http://www.cs.chalmers.se/~aarne/GF/!.

\bibitem{camlp4} Daniel de Rauglaudre. ``The Camlp4 preprocessor."
See: \raggedright \verb!http://caml.inria.fr/camlp4/!.

\bibitem{razet05}
Beno{\^\i}t Razet. Automates modulaires. M\'emoire de Master, 
Universit\'e {Denis} {Diderot} (Paris 7), 2005.

\bibitem{Razet08a}
Beno{\^\i}t Razet. Finite {Eilenberg} Machines.
Proceedings of CIIA 2008, 
Eds. O.H. Ibarra and B. Ravikumar,
Springer-Verlag LNCS vol. 5148 (2008), pp. 242--251.

\bibitem{Razet08b}
Beno{\^\i}t Razet. 
Simulating Finite {Eilenberg} Machines with a Reactive Engine.
In Proceedings of MSFP 2008,
Electric Notes in Theoretical Computer Science,
\url!http://gallium.inria.fr/~razet/PDF/razet_msfp08.pdf!.

\bibitem{Razet09}
Beno{\^\i}t Razet.
Machines d'{Eilenberg} Effectives. Th\`ese de Doctorat,
Universit\'e {Denis} {Diderot} (Paris 7), 2009.

\bibitem{revuz} Dominique Revuz. ``Dictionnaires et lexiques.'' Th\`ese de
Doctorat, Universit\'e Paris VII, Feb. 1991.

\bibitem{rs1} Emmanuel Roche and Yves Schabes. ``Deterministic Part-of-Speech
Tagging with Finite-State Transducers.'' 
Computational Linguistics 21,2 (1995), pp. 227-253.

\bibitem{rs2} Emmanuel Roche and Yves Schabes, Eds.
``Finite-State Language Processing.'' MIT Press, 1997.

\bibitem{sproat} Richard Sproat. ``Morphology and Computation."
MIT Press, 1992.

\bibitem{sproatshih} Richard Sproat, Chilin Shih, William Gale and Nancy Chang.
``A Stochastic Finite-State Word-Segmentation Algorithm for Chinese.''
Computational Linguistics 22,3 (1996), pp. 377--408.

\bibitem{caml} Pierre Weis and Xavier Leroy. ``Le langage Caml.'' 
2\`eme \'edition, Dunod, Paris, 1999. 

\end{thebibliography}

